---
title: "Class 13"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Download and clean HIV-Pr receptor structure
First we download the 1HSG PDB structure from the online database, remove the bound drug from Merck and separate into two new files - one for protein only and one for ligand only. 
```{r}
library(bio3d)
hsg <- get.pdb("1hsg")
hiv <- read.pdb(hsg)
hiv
```
>Q1. What is name of two non-protein resid values in the structure? What does resid correspond to and how would you get a listing of all resid values in this structure?



File we downloaded (1hsg.pdb) contains protein, ligand, and water atoms. So we have to extract the protein atoms only.

Select the protein atoms only with atom.select(). trim.pdb() will perform a similar function. Do the same for the ligand atoms only. Use write.pdb() function to make new files with just these atoms.
```{r}
prot <- atom.select(hiv, "protein", value = TRUE)
write.pdb(prot, file = "1hsg_prot.pdb")
lig <- atom.select(hiv, "ligand", value = TRUE)
write.pdb(lig, file = "1hsg_lig.pdb")
prot
lig
```




